Estimating the number of clusters using Cross Validation

### 2015-06-19 cvclust.tex
add kmeans separation lemma

>>>>

\begin{lemma}
Set $\mu = \int x P(dx)$.
If $\int \|x\|^2 P(dx) < \infty$, then there exists an $r > 0$ such that
for every $A \in \sbA_k$, there is at most one center $a \in A$ with
$\|a - \mu\| \leq r$.
\end{lemma}
\begin{proof}
If the support of $P$ has fewer than $k$ points, then the result is trivial.
Otherwise, Lemma~\ref{lem:kmeans-existence} implies that
$m_k(P) < m_{k-1}(P)$.  Define
\[
  d =
    \inf_{A \in \sA_k}
      \min \{ \|a - a'\| : a,a' \in A, a \neq a' \}.
\]
If $d = 0$, then there exists a sequence $A_n$ in $\sbA_k$ with
\[
  \min_{\substack{a,a' \in A_n \\ a \neq a'}} \|a - a'\| \to 0.
\]
Since $\sbA_k$ is compact, there exists a subsequence $n(m)$ and a set $A_0
\in \sbA_k$ with $A_{n(m)} \to A_0$.
Take $B_n \in \sA_{k-1}$ to be the set formed by removing one of the centers
$a'$ minimizing $\min_{\substack{a \in A_n \\ a \neq a'}} \|a - a'\|$.
By construction, $H(A_n, B_n) \to 0$, so that $B_{n(m)} \to A_0$.
Since the map $A \mapsto W(A, P)$ is continuous on $\sA_k$, we must have
$W(B_{n(m)}, P) \to W(A_0, P) = m_k(P)$.  However, this is impossible since
$W(B_n,P) \geq m_{k-1}(P) > m_k(P)$.  It must be the case, then, that $d > 0$.

Set $r < d/2$, and let $A \in \sbA_k$ be arbitrary.  Take $a_1$ to be the
cluster center in $A$ closest to $\mu$ (breaking ties arbitrarily).  If
$\|a_1 - \mu\| > r$, then $\|a - \mu\| > r$ for all centers $a \in A$.
Otherwise, if $\|a_1 - \mu\| \leq r$ and $a \in A$ is any other cluster center, then
\(
  2r < \|a - a_1\| \leq \|a - \mu\| + \|a_1 - \mu\|
\)
so that
\(
  \|a - \mu\| > r.
\)
\end{proof}
